\thetitle

### 2020-11-19 writing/mutate.tex
revising sample

\begin{figure*}
  \begin{tabular}{p{2.1cm} | p{5cm}  | p{5.5cm} }
    {\bf name} & {\bf description} & {\bf example} \\
\hline

    \texttt{constant}
& Swap a literal constant with another of the same value but different type.
& \texttt{5.6} $\rightarrow$ \texttt{5.6+0.0i} \\
\hline

@@

    \texttt{top-level-id}
 & Swaps two identifiers defined in the same module.
 &\texttt{(f x 42)} $\rightarrow$ \texttt{(g x 42)}\\
\hline

    \texttt{imported-id}
 & Swaps two identifiers imported from the same module.
 &\texttt{(f x 42)} $\rightarrow$ \texttt{(g x 42)}\\
\hline

@@

    \texttt{field-id}
 & Swaps two field identifiers.
    & \begin{tabular}[t]{@{}l} \texttt{(get-field o f)} $\rightarrow$ \\
        \texttt{(get-field o g)}
      \end{tabular}\\
\hline

@@

     \texttt{class:public}
    & Makes a public method private and vice versa.
    & \begin{tabular}[t]{@{}l}
      \texttt{(class object\%}\\
      \phantom{(d}\texttt{(define/public (m x) x))} $\rightarrow$\\
      \texttt{(class object\%}\\
      \phantom{(d}\texttt{(define/private (m x) x))}\\
    \end{tabular}\\
\hline

     \texttt{class:super}
    & Removes \texttt{super-new} calls from class definitions.
    &  \begin{tabular}[t]{@{}l}
      \texttt{(class foo\% (super-new))} $\rightarrow$\\
      \texttt{(class foo\% (void))}\\
    \end{tabular}\\
\hline

@@

     \texttt{negate-cond}
    & Negates conditional test expressions.
    &  \begin{tabular}[t]{@{}l}
      \texttt{(if (= x 0) t e)} $\rightarrow$\\
      \texttt{(if (not (= x 0)) t e)}
       \end{tabular}\\
\hline

     \texttt{force-cond}
    & Replaces conditional test expressions with \texttt{\#t}.
    &  \begin{tabular}[t]{@{}l}
      \texttt{(if (= x 0) t e)} $\rightarrow$\\
      \texttt{(if \#t t e)}
    \end{tabular}\\
\hline
\end{tabular}
  \caption{Summary of mutators.}
  \label{table:mutation-ops}
\end{figure*}

>>>>

\begin{figure*}
  \begin{tabular}{p{2.1cm} | p{7cm}  | p{7.5cm} }
    {\bf name} & {\bf description} & {\bf example} \\
\hline

    \texttt{constant}
& Swaps a constant with another of the same value but different type.
& \texttt{5.6} $\rightarrow$ \texttt{5.6+0.0i} \\
\hline

@@

    \texttt{top-level-id}
 & Swaps identifiers defined in the same module.
 &\texttt{(f x 42)} $\rightarrow$ \texttt{(g x 42)}\\
\hline

    \texttt{imported-id}
 & Swaps identifiers imported from the same module.
 &\texttt{(f x 42)} $\rightarrow$ \texttt{(g x 42)}\\
\hline

@@

    \texttt{field-id}
 & Swaps two field identifiers.
    &  \texttt{(get-field o f)} $\rightarrow$  \texttt{(get-field o g)}\\
\hline

@@

     \texttt{class:public}
    & Makes a public method private and vice versa.
    & \begin{tabular}[t]{@{}l}
      \texttt{(class object (define/public (m x) x))} $\rightarrow$\\
      \texttt{(class object (define/private (m x) x))}\\
    \end{tabular}\\
\hline

     \texttt{class:super}
    & Removes \texttt{super-new} calls from class definitions.
    & \begin{tabular}[t]{@{}l}
      \texttt{(class foo\% (super-new))} $\rightarrow$ \\
      \texttt{(class foo\% (void))}\\
       \end{tabular}\\
\hline

@@

     \texttt{negate-cond}
    & Negates conditional test expressions.
    &  \texttt{(if (= x 0) t e)} $\rightarrow$ \texttt{(if (not (= x 0)) t e)}\\
\hline

     \texttt{force-cond}
    & Replaces conditional test expressions with \texttt{\#t}.
    & \texttt{(if (= x 0) t e)} $\rightarrow$ \texttt{(if \#t t e)}
\end{tabular}
  \caption{Summary of mutators.}
  \label{table:mutation-ops}
\end{figure*}

### 2020-11-19 writing/rational.tex
revising sample

Equipped with these preliminary definitions, we can precisely define what it means
for blame to add value to a gradual type system for a specific
type-level mistake independently of a program's type annotations. 
\begin{quote}
\it
Given a program $\system$, the natural blame mode is\\
(1) \emph{always useful} iff $\lattice{\system}$ contains no failing natural blame trails.\\
(2) \emph{sometimes useful} iff $\lattice{\system}$ contains successful natural blame trails.\\
(3) \emph{never useful} iff $\lattice{\system}$ contains no successful natural blame
trails.\\
\end{quote}

@@

Trails and their properties give us the tools for a rigorous examination
of blame for natural, transient and erasure Typed Racket in the setting of the 
mutants from section~\ref{sec:mutate}. In line with the discussion so far, 
our experimental process collects data to answer three initial questions for each mutant:
\begin{itemize}
\item[$Q_1$] Is blame valuable for natural?

\item[$Q_2$] Is first blame valuable for transient?

\item[$Q_3$] Is last blame valuable for transient?

@@

Furthermore, our experiment can compare the relative value of blame
between any two of the three semantics. 
\begin{itemize}
\item[$Q_*$] Is blame for X more valuable than blame for Y where X, Y is natural,
  erasure or natural?
\end{itemize}

@@

In detail we answer $Q_1$ for an ill-typed mutant from
section~\ref{sec:mutate} by analyzing the natural blame and
natural exception trails in the scenario lattice of the mutant. A
log reports on success and failure endings.  The first step is to construct the
mutant's scenario lattice and select one of the interesting debugging
scenarios from section~\ref{sec:mutate}.
Our test
bed extends these trails according to the natural-blame programmer.  If no scenarios
can be added to the trail, we check if
the last scenario of the trail type checks or not. If it does, this natural-blame
trail is successful, otherwise it is failing. We repeat the same process for the
same roots once more following the natural-exception programmer. 
Next, based on the collected data, we calculate the
success/failure results for the trails to determine whether blame
is more useful than exceptions for this mutant. Finally, we
obtain a positive answer for
$Q_1$ when the percentage of mutants where blame is more useful
than exceptions is greater than the percentage of mutants where exceptions
are more useful
than blame.  The process is analogous for $Q_2$ and $Q_3$  employing the respective
modes of the rational programmer. For $Q_*$, we also proceed in a similar
manner except that we use modes 
of the rational programmer from different semantics. For instance to
decide whether blame for natural is more valuable than blame for
transient, we compare the usefulness of
the natural blame mode with that of the transient first and transient last
modes for all our mutants. 
Figure~\ref{fig:process} summarizes the
experimental process for one mode of the rational programmer and connects
it with the mutations from section~\ref{sec:mutate}.

>>>>

Equipped with these preliminary definitions, we can precisely define what it means
for blame to useful in the context of a gradual type system for a specific
type-level mistake independently of a program's type annotations. 
\begin{quote}
\it
Given a program $\system$, the natural blame mode is\\
(1) \emph{always useful} iff $\lattice{\system}$ contains no failing natural blame trails.\\
(2) \emph{sometimes useful} iff $\lattice{\system}$ contains successful natural blame trails.\\
(3) \emph{never useful} iff $\lattice{\system}$ contains no successful natural blame
trails.\\
\end{quote}

@@

Trails and their properties give us the tools for a rigorous examination
of blame for natural, transient and erasure Typed Racket in the setting of the 
mutants from section~\ref{sec:mutate}. In line with the discussion so far, 
our experimental process collects data to answer three initial questions for each mutant:
\begin{itemize}
\item[$Q_1$] Is blame useful in the context of natural?

\item[$Q_2$] Is first blame useful in the context of transient?

\item[$Q_3$] Is last blame useful in the context of transient?

@@

Furthermore, our experiment can compare the relative usefulness of blame
between any two of the three semantics. 
\begin{itemize}
\item[$Q_*$] Is blame for X more useful than blame for Y where X, Y is natural,
  erasure or natural?
\end{itemize}

@@

In detail we answer $Q_1$ for an ill-typed mutant from
section~\ref{sec:mutate} by analyzing the natural blame and
natural exception trails in the scenario lattice of the mutant. A
log reports on success and failure endings.  The first step is to construct the
mutant's scenario lattice and select one of the interesting debugging
scenarios from section~\ref{sec:mutate}.
Our test
bed extends these trails according to the natural-blame programmer.  If no scenarios
can be added to the trail, we check if
the last scenario of the trail type checks or not. If it does, this natural-blame
trail is successful, otherwise it is failing. We repeat the same process for the
same roots once more following the natural-exception programmer. 
Next, based on the collected data, we calculate the
success/failure results for the trails to determine whether blame
is more useful than exceptions for this mutant. Finally, we
obtain a positive answer for
$Q_1$ when the percentage of mutants where blame is more useful
than exceptions is greater than the percentage of mutants where exceptions
are more useful
than blame.  The process is analogous for $Q_2$ and $Q_3$  employing the respective
modes of the rational programmer. For $Q_*$, we also proceed in a similar
manner except that we use modes 
of the rational programmer from different semantics. For instance to
decide whether blame for natural more useful than blame for
transient, we compare the usefulness of
the natural blame mode with that of the transient first and transient last
modes for all our mutants. 
Figure~\ref{fig:process} summarizes the
experimental process for one mode of the rational programmer and connects
it with the mutations from section~\ref{sec:mutate}.

### 2020-11-19 writing/sample.tex
revising sample

\section{How to Sample the Experimental Space} \label{sec:sample}

As we mention above we treat each definitions as a component. A
consequence of this decision is that examining the whole lattice
for each of the mutants becomes unfeasible.
As figure~\ref{table:benchmarks} shows,  most of the programs contain dozens of definitions  
resulting in lattices with as many as $1.41 \times 10^{28}$ configurations.
Since running a single configuration can take 10 minutes or more 
(due to contract checking and instrumentation), running every possible configuration is impractical.
Evidence from similar contexts suggests that randomly sampling the lattice instead of
an exhaustive exploration is an effective alternative~\cite{greenman2019evaluate}.
Hence we randomly sample enough significant configurations
from the lattice to obtain a confidence of 0.95 (with margin of error
0.05) about whether our hypotheses hold or not.
See accompanying appendix for the details of the calculations for the
sufficient size of samples for this estimate.
Exploring more of the lattice would yield higher confidence in the generalizability
of our results for each benchmark in cases where no violations of a property
(e.g. Blame Trail) are found.
In such cases, we could be more certain that we did not simply miss
the configuration(s) that reveal the violation by exploring the lattice more fully.
Our choice of random sample counts reflects the (informal) standard practice of estimating
results to 95 percent or higher confidence.

>>>>

\section{How to Sample the Experimental Space}

The mutators for section~\ref{sec:mutate} generate 96,515,744 debugging scenarios; 
far too many to even identify the interesting ones among
them. Hence, we perform stratified uniform random sampling with two
levels of stratification to collect sufficiently many interesting
debugging scenarios for drawing statistically sound conclusions. 
The outcome of our sampling process is that we test each mode of the
rational programmer on 72192 randomly selected interesting debugging
scenarios in total spread across both benchmarks and mutators.

As outlined in section~\ref{sec:rational}, all our experimental questions
require that we decide whether a mode of the rational programmer is
always, sometimes or never useful for all the interesting scenarios of a
mutant $\system$. Assuming a normal distribution for the 
interesting scenarios in the lattice $\lattice{\system}$, we can obtain an
estimate answer with high confidence by randomly picking a sufficient
number of interesting scenarios $\lattice{\system}$. Thus
$\lattice{\system}$ becomes the first stratum for our sampling. For
95\% confidence, that is a 5\% margin, the sample size needs
to be just 95 out of $2^N$ scenarios in $\lattice{\system}$ where $N$ is the
number of components of $\system$.

The second sampling is the set of mutants. After all, our mutators
generate 16,800 mutants with
interesting scenarios; still far too many to explore even after sampling
the lattice of each mutant.  Unlike the lattice of scenarios of a mutant
, the set of mutants is not a coherent and finite population;  it is 
unclear how we can relate a sample of mutants to the
of all possible mutants. Thus we sacrifice the ability to generalize 
beyond the mutants we randomly select and opt for the largest random
sample that keeps the experiment computationally feasible. 
In particular,  we sample 80 mutants with interesting scenarios
per benchmark and for any benchmark, we ensure that our sample reflects the diversity of mutants with respect
to the mutators that generated them. Specifically, we select the 80 mutants evenly-distributed
across all of the mutators that contribute mutants for a benchmark.

%% we use the following algorithm to select the 80 mutants evenly distributed across all of the mutators that do produce mutants of a benchmark:
%% 1. Treat each mutator as a bucket containing the mutants produced by that mutator
%% 2. Let the target-sample-size be 80
%% 3. Let the bucket-sample-size be target-sample-size/number-of-buckets
%% 4. For each bucket, do
%%    1. Randomly remove min(bucket-sample-size, mutants-in(bucket)) mutants from the bucket
%%    2. Set target-sample-size = target-sample-size - (the number of mutants removed)
%% 5. If target-sample-size is empty, done; else, go back to 3.
